Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-28 paper.qmd
Added UMAP first choose initially

@@

nldr_two_curvy_original <- umap_two_curvy_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos)
```

@@

trimesh_removed_two_curvy <- vis_rmlg_mesh(
  distance_edges = distance_two_curvy, 
  benchmark_value = benchmark_two_curvy, 
  tr_coord_df = tr_from_to_df_two_curvy, 
  distance_col = "distance") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("d", sc_ltr_pos) 
```

@@

```{r}
#| label: two-curvy-true-proj

true_model_two_curvy <- true_model_df |> 
    select(-ID) |> 
  mutate(type = "true model")

# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)

@@

trimesh_two_curvy_removed1_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
            alpha = 0.5,
            size = 0.5) +
  coord_equal() +
  interior_annotation("b")

@@

nldr_two_curvy_original + 
  trimesh_two_curvy_removed1_with_data +
  two_curvy_proj_umap_model +
  two_curvy_proj_umap_all_model +
  plot_layout(ncol = 2)
```

>>>>

```{r}
#| label: two-curvy-true-proj-data

true_model_two_curvy <- true_model_df |> 
    select(-ID) |> 
  mutate(type = "true model")
```

@@

nldr_two_curvy_original <- umap_two_curvy_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, color = "#636363", size = 0.5) +
  xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
  interior_annotation("a", sc_ltr_pos)
```

@@

trimesh_removed_two_curvy <- vis_rmlg_mesh(
  distance_edges = distance_two_curvy, 
  benchmark_value = benchmark_two_curvy, 
  tr_coord_df = tr_from_to_df_two_curvy, 
  distance_col = "distance") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("d", sc_ltr_pos)

tr_from_to_df_two_curvy <- inner_join(
  tr_from_to_df_two_curvy, distance_df_small_edges_two_curvy,
  by = c("from", "to"))

trimesh_two_curvy_removed1_umap_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
             alpha = 0.5,
             size = 0.5) +
  coord_equal() +
  interior_annotation("b")
```

```{r}
#| label: best-fit-two-curvy-proj-umap1

df_bin_two_curvy <- df_bin_two_curvy |>
  select(-hb_id) |>
  mutate(type = "model")

# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy, df_bin_two_curvy)

model_df <- dplyr::left_join(
  distance_df_small_edges_two_curvy |> select(-distance), 
  projected_model_df, 
  by = c("from" = "ID"))

two_curvy_proj_umap_first_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("c") +
  theme(
    legend.position = "none"
  )

two_curvy_proj_umap_all_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("d") +
  theme(
    legend.position = "none"
  )

@@

```{r}
#| label: two-curvy-true-proj
# Apply the scaling
df_model_data_two_curvy <- bind_rows(data_two_curvy, true_model_two_curvy)

@@

trimesh_two_curvy_removed1_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy3,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = umap_two_curvy_scaled,
             aes(
               x = UMAP1,
               y = UMAP2
             ),
             color = "#636363",
            alpha = 0.5,
            size = 0.5
            ) +
  coord_equal() +
  xlim(c(-0.01, 1.1)) + ylim(c(-0.08, 0.57)) +
  interior_annotation("b")

@@

nldr_two_curvy_original + 
  trimesh_two_curvy_removed1_umap_with_data +
  two_curvy_proj_umap_first_model1 +
  two_curvy_proj_umap_all_model1 +
  plot_layout(ncol = 2)
```
